Natural sources of vagueness and their implications

### 2017-02-20 book_chapter/paper.tex
Wrote about Skyrms and the use of evolutionary game theory for signaling games.

Brian Skyrms~\parencite*[80--104]{skyrms_evolution_1996} identifies some problems with the story so far.
Lewis' account of the stability of conventions rests on what could be considered strong demands for there to be the required common knowledge between the players involved.
Namely, there needs to be a state of affairs that indicates to everyone involved that a certain regularity will hold, as well as ``mutual ascription of some common inductive standards and background information, rationality, mutual ascription of rationality, and so on''~\parencite*[56--57]{lewis_convention_1969}.
These requirements can seem even more excessive if we consider how simple signaling systems are, compared to human languages.
The purpose of studying them was to tell a story about how language as a conventional system could get off the ground without any prior sort of agreement.
It is implausible to assume a high degree of rationality of the agents that started making use of primordial signaling systems that (hypothetically) evolved into languages.
Also, communication through simple message exchange is something that almost all animals do, from monkeys to birds to bees.
A plausible account of the origin of language should first explain how signaling systems like those could get started without a great deal of rationality from the part of the agents involved.

>>>>

The approach adumbrated so far is not, nor does it attempt to be, a full-blown theory of meaning like the classical picture.
However, we can already see it attempts to address the study of meaning from a very different angle.
There is a focus, not on meaning as a kind of correspondence, but rather on the use of signals for specific purposes.
There is also no appeal to truth as a guide for our inquiry.
In this scenario, agents are assumed to possess and exercise rationality.
It is not, however, the same rationality as we find in the classical picture, \emph{i.e.}~the willingness to abide by truth; it is rather a more pragmatic notion of rationality as utility maximization.
One advantage is that we are no longer tied to the need to explain how language hooks on to the world, since this is not relevant on this account.
We are merely trying to see how agents can use signals to cope with the world and achieve their purposes.
On the way, we will better understand meaning, not by trying to say what meaning is, but by seeing how language works.

Brian Skyrms~\parencite*[80--104]{skyrms_evolution_1996}, however, identifies some problems with the story so far.
Lewis' account of the stability of conventions rests on what could be considered strong demands for there to be the required common knowledge between the players involved.
Namely, there needs to be a state of affairs that indicates to everyone involved that a certain regularity will hold, as well as ``mutual ascription of some common inductive standards and background information, rationality, mutual ascription of rationality, and so on''~\parencite*[56--57]{lewis_convention_1969}.
These requirements can seem excessive, even more so if we consider how simple signaling systems are, compared to human languages.
The models were introduced in order to help explain how language as a conventional system could get off the ground without any prior sort of agreement.
If we think of the origins of language in the history of mankind, it seems implausible to assume a high degree of rationality of the agents that started making use of primordial signaling systems that (hypothetically) evolved into languages.
Also, communication through simple message exchange is something that almost all animals do, from monkeys with their calls, to birds with their songs, to bees with their dances, to ants with their pheromone trails.
A plausible account of the origin of language should first explain how signaling systems like those could get started without a great deal of rationality from the part of the agents involved.

In order to address this problem, Skyrms proposes we study signaling problems in evolutionary terms.
Rather than imagining, as Lewis does, rational agents making conscious decisions in possession of knowledge of the game and expectations of the behavior of other agents, we can imagine a simpler scenario inspired by biological evolution: there is a population of agents with biologically hardwired behaviors for engaging in interactions characteristic of a signaling problem; utility does not represent preference, but rather fitness for survival and reproduction; the make-up of the population evolves based on the relative fitness of the strategies represented in the population.
Such a setup attempts to capture the main features of natural selection: in a diverse population, agents with more successful strategies thrive, while agents with less fit strategies die off.
Although the inspiration for this scenario is biological evolution, similar things could be said~\parencite[\emph{e.g.}][]{dawkins_selfish_1978,boyd_culture_1985} about how ideas spread in a population of agents who can adopt or abandon them depending on how successful they appear to be.
The principles can be captured in a formal model that abstracts away from the interpretations: the replicator dynamics.
The only thing relevant to this equation are the relative proportions of strategies in a given population and the utility function.
Using it, one can compute which strategies evolve under which conditions.

The characterization of signaling problems in terms of evolutionary game theory allow us to explain why certain equilibria come to be and how.
Not only can we better understand why signaling systems are stable even without any assumptions of rationality, but we can also map out which initial conditions drive the system towards which equilibria and which don't\todo{mention something about phase portraits}.
In a simple case like the example discussed above, an evolutionary process always drives the system into a state where one signaling system takes over the whole population; which one of the possible two will depend on their relative proportions in the original population.
More complex signaling problems may have different evolutionary outcomes, sometimes surprising ones.
Skyrms~\parencite*{skyrms_signals_2010} gives an overview of different topics studied using signaling games, including expansions of the framework itself (for example, considering other dynamics beyond the replicator equation), exploration of other factors that impact the evolution of signaling (for example, how agents are interconnected), or variations on the signaling problem and its basic assumptions (for example, explanations of deception).

Skyrms' evolutionary game theory approach not only gives more plausible grounds to support Lewis' discussion of convention, but it also accomplishes an important conceptual change, in that it moves most of the theory and mathematical formalisms to the descriptive side of the investigation.
Utility represents how the modeler views the signaling problem and understands the relative advantages or disadvantages of different possible strategy combinations.
Dynamics describe how strategies can evolve when driven by mechanisms of utility maximization.
Focus is put on understanding how various ingredients to the model interact, and which results they produce, not on metaphysical concerns.
While the general framework manages to abstract some details away from the formalization, it does leave room for them, especially when it comes to the dynamics.
We already mentioned the replicator equation that can be seen as representing biological or cultural evolution, but one can also use dynamics inspired on learning mechanisms, or even ones assuming a high degree of knowledge of the game and other players.\todo{give examples and references?}
This range of options goes hand in hand with assumptions of rationality, from no assumptions in a biologically-inspired setting, to a certain degree of rationality captured by a learning method, to higher level rationality and even game-theoretical reasoning.

\todo[inline]{How about vagueness?}
